\section{Understanding the Impacts of Health Spending on Health Outcomes}\label{sec:results}
\setstretch{1.5}

The goal of this section is to understand the impacts of health spending on health outcomes. For that, we first present the estimates of the impact of EC/29 on fiscal and spending outcomes. Later, we analyse how health expenditure increases translate into health inputs. Lastly but not least, we examine the impacts on infant mortality rates.

\subsection{Municipalities' Fiscal Response to the EC/29}

First, we analyse municipalities' fiscal reaction to the EC/29 using data from Finbra\footnote{In section \ref{sec:data} we describe the different sources of public spending data.}. Table \ref{table:fiscal} shows the estimates for total public revenue, total public spending (by type and category), public health spending (by source and type).  Column 1 presents our baseline estimates, a continuous DiD with municipality and state-year fixed effects. Column 2 adds to the baseline specification baseline controls interacted with time. Column 3 adds socioeconomic time varying controls, and column 4, our preferred specification, adds time varying fiscal controls. 

Panel A shows no significant effect on total revenues per capita. Next, in Panel B, we present the effects on total public spending per capita, public spending by type and category. We find no significant effect on total spending, nor on public spending by type. When it comes to spending by category, the distance to the EC/29 target is only significantly associated with Health and Sanitation spending per capita\footnote{This is the only measure of Health Spending available for the period before the EC/29 was enacted. As discussed in section \ref{sec:emp}, the validation of our research design relies partially on evaluating the presence of pre-trends. Therefore we use this variable as a proxy of health spending. In the appendix we show that}. This result is extremely important for two main reasons reasons. First, it validates the descriptive evidence showing that EC/29 was responsible for increasing municipalities' health expenditure. Second, it allow us to claim that reductions in Infant Mortality rates are associated with increases in health expenditure and not increases in spending from other categories that could also affect mortality, such as social assistance and education. In our preferred specification (column 4), the estimate of around 300 suggests that, a municipality with a distance to the target of 10\%, increased its health spending per capita by R\$30, an increase of around 14\% relative to the baseline health spending per capita (see Table \ref{table:stats}). This distance is roughly the distance to the target of the bottom quartile of the distribution of the share of own resources spent in health in the baseline. 

Figures \ref{fig:6} and \ref{fig:7} show flat and insignificant effect on revenues, total spending and spending by type throughout the years after the constitutional amendment. Figure \ref{fig:8} plots the trends for public spending by category. Data for some of this variables is quite noisy for the period before the year 2000, but in general, for all categories but Health and Sanitation, we find flat and insignificant estimates through the years. Figure \ref{fig:8a} shows a clear and significant increase in Health and Sanitation spending per capita. For each of the first years after the EC/29 the effect becomes stronger, stabilizing from 2005 on.

\input{tables/Table_1_fiscal}


Though SIOPS data is only available from 2000 on, it allow us to analyse the sources of spending and the types of spending within total public health spending. Panel C in Table \ref{table:fiscal} presents the results for these variables. These results should be taken with parsimony, as these regressions do not include data for pre-treatment periods. Our estimations suggests an effect of R\$ 528 for total Health Spending per capita. A distance of 10\% to the target represents a 27\% increase relative to the baseline, almost twice the effect on Health and Sanitation per capita. It makes sense that a variable that only includes health spending presents a higher effect, as our source of variation is expected to move health spending only. Additionally, as presented in Panel C this effect comes almost entirely from increases in spending from own resources, a five fold increase relative to own resource spending in the baseline. We also find some substitution effects, with municipalities reducing some of its spending from intergovernmental transfers in health. All types of health spending were responsible for this increase in total health spending, but the increase in investment is the noteworthy, specially in relative terms. A distance of 10\% is associated with a 90\% increase in health investments. Baseline statistics show very little resources allocated in investments within total public health spending, the great majority of resources were allocated in human resources and in other administrative expenses. Considering the importance of capital investments to the supply of medical resources and the quality of medical services, and the little amount of investments in the baseline, an effect of this size is really relevant.

Though at different levels, the effects for SIOPS total health spending follow a very similar dynamic pattern as the effects for Finbra health and sanitation spending. These patterns are presented side by side in Figure \ref{fig:9a} and \ref{fig:9b}. In Figure \ref{fig:10} 

[DYNAMIC EFFECTS]

Figures \ref{fig:9} \ref{fig:10}

\input{figures/Figure_06}


\input{figures/Figure_07}


\input{figures/Figure_08}


\input{figures/Figure_09}


\input{figures/Figure_10}


\subsection{Effects on Health Inputs}

In this subsection we aim to explore the pathways that mediate the relationship between health spending and health outcomes. For that, we explore the impacts of the EC/29 on several health inputs: access to public health services, health infrastructure, and ambulatorial production.

Table present the estimates for several health infrastructure proxies. For the period of analysis there is no direct health infrastructure data available. Using the SIA data, we are able, for each year, to count the number of facilities with ambulatorial services, in general, and classified by the type of service and type of professional associated with the service. 

Table \ref{table:infra}, Figures \ref{fig:11} \ref{fig:12} \ref{fig:13}

\input{tables/Table_2_coverage_infra}

\input{figures/Figure_11}

\input{figures/Figure_12}

\input{figures/Figure_13}

Table \ref{table:production}, Figures \ref{fig:15} \ref{fig:14}

\input{tables/Table_3_access_production}

\input{figures/Figure_14}

\input{figures/Figure_15}



\subsection{Effects on Infant Mortality}

Having provided evidence of meaningful increases in health spending, we now present estimates of the causal effects on infant mortality. Differently from most of the literature linking health spending with infant mortality \citep{filmer1999,bokhari2007,moreno2015,nixon2006,gupta2002effectiveness,cremieux1999,bokhari2007}, we are able to assess the effects not only for total infant mortality rates, but also for rates by timing of death, rates by cause of death, and also infant mortality rates that are amenable to primary care. These results are presented in Table \ref{table:imr}. In our preferred specification, with the exception of ill-defined infant mortality, all outcomes present negative point estimates, but only infant mortality rate within 24 hours and infant mortality rate by infectious diseases present marginally significant estimates. Relative to the baseline mortality rates, a 10\% distance to the target is associated with a reduction of around 3.5\% for mortality within 24 hours and a reduction of around 5\% for mortality due to infectious diseases. 

Though most of the continuous DiD estimates for Infant Mortality Rates presented in Table \ref{table:imr} are not statistically significant, the more flexible coefficients estimated with equation \ref{eq:2} provide useful information on the dynamics of the effects and suggest the presence of some significant reduction in infant mortality rates. The estimates of these regressions and its confidence intervals of 5\% are presented in Figures \ref{fig:imr1}, \ref{fig:imr2} and \ref{fig:imr3}. Total infant mortality rate (\ref{fig:imr1_a}) and infant mortality rate amenable to primary care (\ref{fig:imr1_b}) present a clear trend of reduction, with estimates for later years being statistically significant. Mortality from 1 to 27 days old (Figure \ref{fig:imr2_c} and from 27 days to one year old (Figure \ref{fig:imr2_d} presents some evidence of reduction through the years after treatment. When it comes to infant mortality rates by cause of death, mortality due to perinatal conditions (Figure \ref{fig:imr3_c}) presents a clear trend of reduction with all years from 2005 on presenting statistically significant estimates.

In general, articles estimating the causal relationship between health spending and mortality run log-log regressions and present estimates for the elasticity of mortality with respect to health spending. We explicitly choose not to apply transformations to our health outcomes variables due to the amount of observations with values equal to 0, notably the ones related to birth and mortality. As we work with a data set comprising all the Brazilian municipalities with available data for the period of analyses, some with population size as little as 700 inhabitants, it is common to find infant mortality rates of 0. Running log transformation would throw away some relevant information for several outcomes. Nonetheless, to relate our results to the literature on this topic we estimate "back of the envelope" elasticities for all mortality rates presented so far. Table \ref{table:elasticity} presents these elasticities for our preferred specification estimates using Finbra Health and Sanitation Spending per capita and Infant Mortality Rates. 

Figures \ref{fig:16} \ref{fig:17} \ref{fig:18}

\input{tables/Table_4_imr}

\input{figures/Figure_16}

\input{figures/Figure_17}

\input{figures/Figure_18}


[Compare with other articles elasticities]


Table \ref{table:birth} presents the estimates for a few Birth Outcomes. In general, the point estimates are in the expected direction, substantially small, and statistically insignificant.      

Figure \ref{fig:19}

\input{tables/Table_5_birth}



