\documentclass[12pt, final]{article}

\input{_preamble}
\usepackage{etoolbox}
\usepackage{tocloft}


%\newcommand\aauthorA{Michel Szklo}
\newcommand\aaffiliationA{FGV-SP}

\begin{document}


\title{\textbf{Public Health Spending and Infant Mortality: Evidence from a Constitutional Reform in Brazil}}
\pagenumbering{gobble}

\author{
\small{Michel Szklo}\footnote{FGV - S\~ao Paulo School of Business Administration. Contact email: \href{mailto:mszklo@gmail.com}{mszklo@gmail.com}.}
\and \small{Damian Clarke}\footnote{Department of Economics, Universidad de Chile, MIPP \& IZA. Address: Diagonal Paraguay, 257, Santiago, Chile.  Contact email: \href{mailto:dclarke@fen.uchile.cl}{dclarke@fen.uchile.cl}.}
\and \small{Rudi Rocha}\footnote{FGV - S\~ao Paulo School of Business Administration, rudi.rocha@fgv.br.}
}
\date{\today}
\maketitle
\begin{center}
    %First draft: July 2020 \\
    %\emph{Preliminary. Please do not circulate}
\end{center}
\singlespacing
\begin{abstract}
 \noindent This paper treats Brazil as a case study to document the causal effects of health spending on infant mortality. By leveraging the variation in health spending promoted by Brazil's \nth{29} Constitutional Amendment of 2000, we are able to document not only the effects of health spending on infant mortality, but also the links in the chain connecting spending to health outcomes. We show that increases in health spending translate into greater primary care coverage, higher supply of hospitals and low skilled professionals. Associated with this effects we document moderate reductions in infant mortality within 24 hours and due to perinatal conditions, as well as long term reductions in total infant mortality, infant mortality amenable to primary care, and infant mortality by infectious and respiratory causes.  Our results contribute to the literature on the impacts of health spending by providing on of the first well identified causal parameter of the relationship between spending and mortality.
\end{abstract}

\noindent \textbf{JEL Codes}: I1, I3, O5\\
\noindent \textbf{Keywords}: Health spending; health care reform; health care production; infant mortality.

\newpage
%\thispagestyle{empty} 
%\onehalfspacing 
\setcounter{page}{1}



\input{1_intro}
\input{2_institutional}
\input{3_data}
\input{4_empirical}
\input{5_results}
\input{6_conclusion.tex}

\clearpage
\pagebreak
\singlespacing  
\bibliographystyle{apalike}
\bibliography{SRC}
\nocite{}




\pagebreak
\appendix
\singlespacing	
\input{appendix}


\end{document} 

